Operational adoption of new probabilistic point rainfall forecasts: the crucial role of user-centric training

### 2020-11-12 manuscript/main.tex
Updates from Overleaf

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TemplateV5.tex --  LaTeX-based template for submissions to the American Meteorological Society
% Version 5.0, 2 January 2020
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Packages
\usepackage{amsmath,amsfonts,amssymb,bm}
\usepackage{mathptmx}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{xcolor}

%%%%%%%%%
% TITLE %
%%%%%%%%%
\title{Improving the communication of new point rainfall forecasts through a dialogue with forecasters}

@@

\extraauthor{Hannah Cloke}
\extraaffil{University of Reading, Reading, UK \\
Uppsala University, Uppsala, Sweden}

%%%%%%%%%%%%
% ABSTRACT %
%%%%%%%%%%%%
\abstract{In April 2019, new statistically post-processed point-rainfall forecasts (ecPoint-Rainfall) became operational at the European Centre for Medium-Range Weather Forecasts. An extensive verification study shows that ecPoint-Rainfall has higher reliability/skill than raw ECMWF ensemble for forecasting rainfall at a point. However, ecPoint-Rainfall adds layers of complexity to standard probabilistic ensemble-based forecasts (PEFs) and, before this study, it was not clear to what extent forecasters might find its interpretation conceptually challenging, even for those with an extensive probabilistic forecasting experience. This study was therefore deemed necessary to evaluate whether traditional guidelines for PEFs would be sufficient for the adoption of ecPoint-Rainfall, or whether they would need to be somehow tailored to take account of the specific features of this new product to best communicate and exploit its added value. During a one-year trial period, ecPoint-Rainfall was tested in an operational environment at two meteorological centres with different expertise on PEFs: the Hungarian Meteorological Service and the National Meteorological Institute of Costa Rica. Later, crucial aspects that conditioned or might condition in future the adoption of ecPoint-Rainfall were explored with the forecasters that used ecPoint-Rainfall in their day-to-day work. This paper summarises the outcomes of the initial independent feedback and the subsequent discussions on how valuable ecPoint-Rainfall was perceived. The overall impression is that ecPoint-Rainfall could improve the prediction of extreme localized rainfall events only if tailored training is provided, no matter what is the prior forecasters experience on PEFs.}

@@

\textit{"Forecasts possess no intrinsic value. They acquire value through their ability to influence the decisions made by users of the forecasts”} \citep{Murphy1993}. With this sentence, Murphy states that increases in forecast accuracy do not necessarily imply increases in forecast value as the quality/value relationship is user and problem-specific. \par

@@

Therefore, even the most recent discussions \citep{Zhang2019} keep highlighting how important it is to give proper attention to the whole forecasting chain to expand the adoption of PEFs to a broader group of end-users. This includes (i) continuous improvements in the formulation of probabilistic weather predictions, (ii) using post-event analysis to constantly evaluate and update communication strategies for low-probability high-impact events without reducing trust through, (iii) user-engagement, (iv) user-testing, and (v) awareness of design principles. \par
It is within this framework that this study proceeds. In 2019, ECMWF made operational a new post-processed point-rainfall product, called ecPoint-Rainfall. A one-year global verification showed that, for points, ecPoint-Rainfall provides more reliable and skilful rainfall forecasts than the raw ECMWF ensemble (ENS), with marked improvements in the discrimination ability for extreme rainfall events (> 50 mm/12h) in the medium range being particularly noteworthy \citep{Hewson2020a}. However, they did not assess the new product from a user perspective, and this constitutes a gap in their study. ecPoint-Rainfall adds layers of complexity to standard PEFs and, without a user-focused evaluation, it is impossible to assess to what extent end-users, including those experienced in probabilistic weather forecasting, might find the interpretation of the new forecasts conceptually challenging. Such a complexity might lead to a misunderstanding or misuse of the product, eventually undermining a possible broader adoption. Therefore, this study aims to assess whether traditional PEF guidelines are appropriate also for ecPoint-Rainfall, or whether they would need to be somehow tailored to best communicate the specific features and the added value of this new product, even in case of experienced users of probabilistic forecasts. \par
Hungary and Costa Rica, that have meteorological services with different experience in probabilistic weather forecasting, volunteered to test ecPoint-Rainfall within their operational environments over a one-year trial period. Both  meteorological services reported on the ecPoint-Rainfall performance in predicting extreme rainfall events that caused significant impacts in their own countries during the trial. Such analysis complements the more general verification carried out by \citet{Hewson2020a} by illustrating the performance of ecPoint-Rainfall in two different climatological regions of the world. Then, subsequent discussions were carried out to understand how ecPoint-Rainfall was viewed  operationally, and which aspects might condition in future its full adoption within operational environments. \par

@@

@@

\begin{figure}
\centerline{\includegraphics[width=19pc]{Figures/Fig1.jpg}}
\caption{Orography of Costa Rica (a) and Hungary (b). The insert box shows the location of Hungary and Costa Rica in Europe and America, respectively.}
\label{Fig1}
\end{figure}

{\color{blue}
\subsection{National Meteorological Institute of Costa Rica (IMN)} 
IMN, in Costa Rica (see Fig.\ref{Fig1}a), has a recent experience with EPSs and with ECMWF forecasts, which started using in 2018. They work mainly with the Global Forecast System (GFS) developed by the National Centers for Environmental Prediction in USA and with the Weather Research and Forecasting (WRF) system. From the latter, IMN has developed a series of model versions (WRF-1.5, WRF-5, WRF-AR, WRF-Sarapiquí, WRF, WRF-8, WRF-15) that are tailored to different needs, such as prediction of tropical waves, tropical cyclones, cold fronts, hail and forest fires. Such versions have different spatial and temporal resolutions, different domains, and different model configurations. The most recent model, WRF-1.5, was developed in 2018 at IMN to produce forecasts at 1.5 km resolution up to 5 days in advance to improve the predictions for convective systems that can generate extreme rainfall events in Costa Rica. 
}

{\color{red}
\subsection{Hungarian Meteorological Service (OMSZ)}
OMSZ, in Hungary (see Fig.\ref{Fig1}b), has a long experience working with EPSs. They are at the forefront of developing ensemble based products, adopted also by other centres, see for example \citet{Gascon2018}, and they use probabilistic forecasts in their day-to-day operational duties to use warning to the public and end-users. OMSZ has a long history of collaborations with ECMWF. Not only OMSZ has a long tradition of carrying out objective verification of ECMWF ensemble forecasts \citep{Hewson2020}, but it also uses widely ECMWF’s software and archived forecasts for all time ranges since 1995. Forecasters from OMSZ also take part regularly in educational and training programmes at ECMWF. \par
OMSZ has also experience in post-processing of rainfall forecasts. Objective verification has shown that calibration can add value to rainfall forecasts for small-scale low-predictability phenomena, such as extreme rainfall events, and such positive calibration results can get reflected also on downstream applications, such as hydrological forecasts \citep{Ihasz2018,Matrai2017}.
}

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{Capture.png}}
\caption{Panel (a) shows the climate regions of Costa Rica defined by IMN. Panels from (o-1) to (o-6) show 12-hourly rainfall observations between October 3rd and 5th, 2018. Panels from (i-1) to (i-3) show the impacts caused by the extreme event in the regions in which CNE declared the state of red alert (the photos come from www.teletica.com).}
\label{Capture}
\end{figure*}

@@

ecPoint is a statistical post-processing technique that provides probabilistic bias-corrected forecasts at point scale from ensemble or deterministic numerical weather prediction (NWP) model outputs. Different hydro-meteorological variables, from different NWP models, can potentially be post-processed using the ecPoint technique, generating the so-called "ecPoint Family". ecPoint-Rainfall is the branch of the family that produces rainfall forecasts that mirror rain gauge measurements \citep{Hewson2020a}. In this paper, the ecPoint-Rainfall forecasts post-process raw ECMWF ENS rainfall forecasts. It is worth noting that what described to the following also applies to other NWP models. \par
To explain the reasons why ecPoint-Rainfall and raw ECMWF ENS rainfall forecasts may differ, let us consider an ENS grid cell and radar-derived totals for a rainfall event. Let us considered that the grid cell average rainfall total is about 17mm, while the minimum and maximum rainfall amounts are about 2mm and 60mm, respectively. The following can be stated. First, a completely accurate ENS member forecast would predict 17mm. However, this would not give the user any idea on what could be the spatial variability, within the grid cell, of the local (i.e. point) rainfall amounts when a particular G\_WT occurs. This describes an issue of sub-grid variability within the model grid box. Secondly, a particular G\_WT might show that an ENS member tends, on average, to over-predict (or under-predict) the grid-box rainfall by 15\%, delivering a forecast of 20mm (or 14mm) instead of 17mm. This describes an issue of model bias at grid-scale. By anticipating the sub-grid variability within the grid-box and correcting for biases in the model at grid-scale, ecPoint-Rainfall provides probabilities of bias-corrected rainfall forecasts at a point within the grid-box. However, ecPoint-Rainfall does not say where that point is within the ENS grid-box. \par
The calibration dataset for ecPoint-Rainfall is built applying the new concept of "Remote Calibration" described in \citet{Hewson2020a}. Typically, traditional post-processing methods use observations from a particular site. Such an approach limits the size of the calibration dataset, and forces to wait for at least ~20 years to have a robust calibration dataset, which still might not contain (enough) extremes cases for the considered site. The "Remote Calibration" approach removes any location labels and, instead, generates calibration datasets for particular G\_WTs pulling observations from where, around the world, those G\_WTs occurred. This approach facilitates the generation of extensive training datasets to post-process rainfall, including extremes. Indeed, the sites where a certain G\_WT occurs relatively often build the calibration dataset also for those sites where the same G\_WT would generate extreme rainfall. To give an order of magnitude, in order to create a similar calibration dataset for a location with similar size of one generated with the "Remote Calibration" approach, one would need to wait from hundreds to thousands of years. It is essential to highlight that the error bars on forecasts for globally extreme rainfall events (e.g. tropical cyclones) will be inevitably larger than for "not globally" extremes due to a lack of events in the calibration dataset. \par
\citet{Hewson2020a} have shown that, at every lead time, up to 10 days, ecPoint-Rainfall outperforms raw ECMWF ENS when predicting localise rainfall events, including also those extreme. \par

@@

\subsection{Hands-on experience, the trial}
According to \citet{Fundel2019}, users need the opportunity to familiarise with the quality of new products in their everyday practice in order to appreciate their entire value in decision-making processes. OMSZ and IMN volunteered to test the ecPoint-Rainfall forecasts in their day-to-day operational routines. The participation of OMSZ and IMN was ideal as the two meteorological services have different degrees of experience with probabilistic forecasts, and reside in two different climatological regions, tropical for IMN and extra-tropical for OMSZ, so that allows the evaluation of ecPoint-Rainfall in different climates. The trial period started in 2018. During the trial, OMSZ and IMN were invited to analyse the ecPoint-Rainfall performance in predicting extreme rainfall events, and finally summarise their results in a report to send to ECMWF at the end of the trial. Given the small volumes of data available, both countries decided to carry out their analysis on independently selected case studies. The ecPoint-Rainfall forecasts were presented to both countries at the beginning of the trial. They were informed of different sources of information about ecPoint-Rainfall on the web, such as newsletters or the ECMWF Forecaster's User Guide \citep{Owens2018} and no further details were provided unless direct requests were submitted.

\subsection{Dialogue forecasters-developers}
The forecasters who monitored the ecPoint-Rainfall forecasts in their day-to-day operational duties were invited to discuss the outcomes of the reports. Forecasters were also invited to discuss the theoretical and practical challenges that emerged when adopting ecPoint-Rainfall in an operational environment. Such an open-ended approach are particularly well suited to explore behavioural responses to innovative forecast products like ecPoint-Rainfall, where the range of possible conceptual framings and the meaningful behaviour of the informants is not quantifiable but must be teased out through careful interpretation \citep{Patton2002}.

@@

\begin{figure}
\centerline{\includegraphics[width=18pc]{Fig2.jpg}}
\caption{Product developed by IMN based on the 85th percentile of ecPoint-Rainfall.}
\label{Fig2}
\end{figure}

\begin{figure*}
\centerline{\includegraphics[width=39pc]{Fig3.jpg}}
\caption{Panel (a) shows the climate regions of Costa Rica defined by IMN. Panels from (o-1) to (o-6) show 12-hourly rainfall observations between October 3rd and 5th, 2018. Panels from (i-1) to (i-3) show the impacts caused by the extreme event in the regions in which CNE declared the state of red alert (the photos come from www.teletica.com).}
\label{Fig3}
\end{figure*}

@@

\begin{figure}
\centerline{\includegraphics[width=19pc]{Fig4.jpg}}
\caption{Panels (a) and (c) show the probabilities of not exceeding 50 mm/12h, and panels (b) and (d) show the 99th percentile, both for ecPoint-Rainfall forecasts. Panels (a) and (b) correspond to the forecast on 2018/10/04 at 00 UTC (t+18,t+30), which correspond to the period between 2018/10/04 12am and 201/10/05 0am (local time). Panels (c) and (d) correspond to the forecast on 2018/10/04 at 12 UTC (t+18,t+30), which correspond to the period between 2018/10/05 0am and 201/10/05 12am (local time).} 
\label{Fig4}
\end{figure}

IMN reported that the 85th percentile was selected to predict extreme events but to not focus only on the \textit{"worst-case scenarios"}. IMN forecasters pointed out, for example, that events around 50 mm/12h can already cause some impacts on the Pacific coast of Costa Rica or in the capital city, San José. Such events are relatively frequent in Costa Rica. Therefore, IMN forecasters thought they would miss such more frequent events if they look only at the highest percentiles. \par
IMN assessment is fair. On the other hand, by looking only at the 85th percentile (which would forecast events that have, on average, a 1 in 7 chance to occur), forecasters would miss the opportunity to be aware of what could be the possible "worst-case scenario". The latter becomes, indeed, a piece of vital information when preparing forecasts for critical decision-making processes in highly uncertain weather scenarios like the one analyzed. \par
During the dialogue with IMN forecasters, it was discussed the possibility to make available to forecasters the probabilities of exceeding a certain threshold. For example, 50 mm/12h is a rainfall event that can generate some impacts in some areas of Costa Rica. Forecasters could display on a map-plot the probabilities of exceeding 50 mm/12h (see mock-up product in Fig.\ref{Fig4}a and c) and identify whether in the concerning areas the probabilities of exceeding 50 mm/12h are higher than a probability threshold previously defined by IMN for which action needs to be taken. At the same time, by looking at the 99th percentile (see mock-up product in Fig.\ref{Fig4}b and d), forecasters could identify those areas at risk of being affected by the "worst-case scenario" (as it represents those events with, on average, a 1 in 100 chance to occur), and provide a richer warning on the areas at risk of highest impacts. \par
When both products (Fig.\ref{Fig4}) were shown to IMN forecasters forecasters to see whether they would have reformulated their forecasts for the case study, they produced a much more aware warning about the areas at risk of highest impacts (like those in https://www.imn.ac.cr/es/web/imn/avisos-meteorologicos). \par
Furthermore, subsequent examination of different extreme rainfall cases around the world has shown that raw ENS and ecPoint-Rainfall’s CDFs tend to cross around the 85th percentile, and ECMWF users might not extract extra information from ecPoint-Rainfall not provided by the raw ENS.

{ \color{blue}
Perhaps it would be nice to include an example of a warning using ecPoint-Rainfall next to the warning that was actually published for the analysed event.}

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{Fig5.jpg}}
\caption{Forecast evolution (from day 7 to day 1) for the rainfall event occurred between October 04th at 12am and October 5th at 0am (see Fig. \ref{Fig3}o-4 to compare the forecasts with the observations). From the left, the first column shows the 85th percentile for ecPoint-Rainfall, the second column shows the 85th percentile for the raw ENS, the third column shows the 99th percentile for ecPoint-Rainfall, the fourth column shows the wettest member of the raw ENS, and the fifth column shows the deterministic forecast from WRF-1.5 provided by IMN (spatial resolution of 1.5 km). The shown raw ECMWF ENS and ecPoint-Rainfall forecasts correspond to runs at 12 UTC, which is the first available run from Europe to IMN forecasters due to time difference (UTC-6) between Europe and America. The displayed WRF-1.5 forecasts correspond to runs at 18 UTC, which is the first run available to IMN forecasters in the morning for their in-house models.}
\label{Fig5}
\end{figure*}

@@

Fig.~\ref{Fig5} shows the forecasts, up to day 7, for the rainfall event between midday October 4th and midnight October 5th (compare with observations in Fig.\ref{Fig3}o-4). The first and the second column in Fig.~\ref{Fig5} show, respectively, the 85th percentile of ecPoint-Rainfall and raw ENS. The maps in these two columns show to the reader what it was previously stated about the similarity of the raw ENS and ecPoint-Rainfall's 85th percentile. The third, the fourth and the fifth column in Fig.~\ref{Fig5} show, respectively, the 99th percentile of ecPoint-Rainfall, the wettest ENS member, and the rainfall accumulations for the deterministic WRF-1.5 model. ecPoint-Rainfall provides from day 7 a more consistent guidance on the location of the areas at most risk of extreme rainfall, i.e. the whole Pacific coast (including the Nicoya peninsula), with a signal that becomes better defined from day 2 and shows forecast values of the same order as those observed. On the other hand, the raw ENS and the WRF-1.5 provide a noisier and jumpier signal. In the raw ENS the Nicoya peninsula is not highlighted in the guidance until day 2, and the rainfall extremes are underestimated of ~50\% also at day 1.

@@

%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS %
%%%%%%%%%%%%%%%%%%%
\acknowledgments
We thank the anonymous reviewers, Florian Pappenberger, David Richardson and Christel Prudhomme who helped to improve the manuscript with their comments.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TemplateV5.tex --  LaTeX-based template for submissions to the American Meteorological Society
% Version 5.0, 2 January 2020
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% Packages
\usepackage{amsmath,amsfonts,amssymb,bm}
\usepackage{mathptmx}
\usepackage{newtxtext}
\usepackage{newtxmath}
\usepackage{enumitem}
\usepackage{multirow}
\usepackage{xcolor}
\usepackage{svg}
\usepackage{amsmath}

%%%%%%%%%
% TITLE %
%%%%%%%%%
\title{Operational adoption of new probabilistic point rainfall forecasts: the crucial role of user-centric training}

@@

\extraauthor{Hannah L. Cloke}
\extraaffil{University of Reading, Reading, UK \\
Uppsala University, Uppsala, Sweden\\
Centre of Natural hazards and Disaster Science, Uppsala, Sweden}

%%%%%%%%%%%%
% ABSTRACT %
%%%%%%%%%%%%
\abstract{Users of Numerical Weather Prediction (NWP) models often require forecasts for specific locations. When they have access only to raw gridded NWP model outputs, forecasters are trained to interpret such model output on the basis of their experience and expertise. Such expert post-processing adds value by correcting raw forecasts for specific weather conditions, but does have limitations in that particularly when new or more complex forecast products are released. Some of this limitations are .
Following the release of new statistically post-processed point-rainfall forecasts (ecPoint-Rainfall) at the European Centre for Medium-Range Weather Forecasts (ECMWF), this study evaluates to what extent traditional guidelines for standard probabilistic ensemble-based forecasts (PEFs) are sufficient for the adoption of ecPoint-Rainfall. During a one-year trial period, ecPoint-Rainfall was tested in an operational environment at two meteorological centres with different expertise on PEFs: the Hungarian Meteorological Service and the National Meteorological Institute of Costa Rica. Later, crucial aspects that conditioned or might condition in future the adoption of ecPoint-Rainfall were explored with the forecasters that used ecPoint-Rainfall in their day-to-day work. This paper summarises the outcomes of the initial independent feedback and the subsequent discussions on how valuable ecPoint-Rainfall was perceived. The overall impression is that ecPoint-Rainfall could improve the prediction of extreme localized rainfall events only if tailored training is provided, no matter what is the prior forecasters experience on PEFs.}

@@

\textit{"Forecasts possess no intrinsic value. They acquire value through their ability to influence the decisions made by users of the forecasts”} \citep{Murphy1993}. Thus increases in forecast accuracy do not necessarily imply increases in forecast value, as the quality/value relationship is user and problem-specific. \par

@@

Therefore, even the most recent discussions \citep{Zhang2019}  highlight the importance of giving proper attention to the whole forecasting chain to expand the adoption of PEFs to a broader group of end-users. This includes (i) continuous improvements in the formulation of probabilistic weather predictions, (ii) using post-event analysis to constantly evaluate and update communication strategies for low-probability high-impact events without reducing trust through, (iii) user-engagement, (iv) user-testing, and (v) awareness of design principles. \par
In 2019, a new post-processed point-rainfall product, called ecPoint-Rainfall became operational at ECMWF. A one-year global verification showed that, for points, ecPoint-Rainfall provides more reliable and skilful rainfall forecasts than the raw ECMWF ensemble (ENS), with marked improvements in the discrimination ability for extreme rainfall events (> 50 mm/12h) in the medium range being particularly noteworthy \citep{Hewson2020a}. However, the new product was not assessed from a user perspective. ecPoint-Rainfall adds layers of complexity to standard PEFs and, without a user-focused evaluation, it is impossible to assess to what extent end-users, including those experienced in probabilistic weather forecasting, might find the interpretation of the new forecasts conceptually challenging. Such complexity might lead to a misunderstanding or misuse of the product, eventually undermining a possible broader adoption. Therefore, this study aims to assess whether traditional PEF guidelines are appropriate also for ecPoint-Rainfall, or whether they would need to be somehow tailored to best communicate the specific features and the added value of this new product, even in case of experienced users of probabilistic forecasts. \par
Hungary and Costa Rica have meteorological services with different experience in probabilistic weather forecasting, and they volunteered to test ecPoint-Rainfall within their operational environments over a one-year trial period. Both  meteorological services reported on the ecPoint-Rainfall performance in predicting extreme rainfall events that caused significant impacts in their own countries during the trial. Such analysis complements the more general verification carried out by \citet{Hewson2020a} by illustrating the performance of ecPoint-Rainfall in two different climatological regions of the world. Then, subsequent discussions were carried out to understand how ecPoint-Rainfall was viewed  operationally, and which aspects might condition in future its full adoption within operational environments. \par

@@

According to \citet{Fundel2019}, users need the opportunity to familiarise with the quality of new products in their everyday practice in order to appreciate their entire value in decision-making processes.

@@

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Fig1.jpg}}
\caption{Orography of Costa Rica (a) and Hungary (b). The insert box shows the location of Hungary and Costa Rica in Europe and America, respectively.}
\label{Fig1}
\end{figure}

\subsection{National Meteorological Institute of Costa Rica (IMN)} 
IMN, in Costa Rica (see Fig.\ref{Fig1}a), has recent experience with EPSs and with ECMWF forecasts, which they started using in 2018. They work mainly with the Global Forecast System (GFS) developed by the National Centers for Environmental Prediction in USA and with the Weather Research and Forecasting (WRF) system. From the latter, IMN has developed a series of model versions (WRF-1.5, WRF-5, WRF-AR, WRF-Sarapiquí, WRF, WRF-8, WRF-15) that are tailored to different needs, such as prediction of tropical waves, tropical cyclones, cold fronts, hail and forest fires. Such versions have different spatial and temporal resolutions, different domains, and different model configurations. The most recent model, WRF-1.5, was developed in 2018 at IMN in order to produce forecasts at 1.5 km resolution up to 5 days in advance which improves the predictions for convective systems that can generate extreme rainfall events in Costa Rica.

\subsection{Hungarian Meteorological Service (OMSZ)}
OMSZ, in Hungary (see Fig.\ref{Fig1}b), has a long experience working with EPSs. They are at the forefront of developing ensemble based products, adopted also by other centres, see for example \citet{Gascon2018}, and they use probabilistic forecasts in their day-to-day operational duties to use warning to the public and end-users. OMSZ has a long history of collaborations with ECMWF. Not only has OMSZ a long tradition of carrying out objective verification of ECMWF ensemble forecasts \citep{Hewson2020}, but has also been using ECMWF software and archived forecasts since 1995. Forecasters from OMSZ also take part regularly in educational and training programmes at ECMWF. \par
OMSZ has also experience in post-processing of rainfall forecasts. Objective verification has shown that calibration can add value to rainfall forecasts for small-scale low-predictability phenomena, such as extreme rainfall events, and such positive calibration results can get reflected also on downstream applications, such as hydrological forecasts \citep{Ihasz2018,Matrai2017}.

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Fig1.png}}
\caption{Comparison of approaches for forecasts adjustments: expert human adjustment vs. ecPoint.}
\label{Fig.2}
\end{figure*}

@@

ecPoint is a statistical post-processing technique that provides probabilistic bias-corrected forecasts at point scale from ensemble or deterministic numerical weather prediction (NWP) model outputs. Different hydro-meteorological variables, from different NWP models, can potentially be post-processed using the ecPoint technique to create, what is known as, the "ecPoint Family". ecPoint-Rainfall is the branch of the family that produces rainfall forecasts that mirror rain gauge measurements \citep{Hewson2020a}. In this paper, the ecPoint-Rainfall forecasts are post-processed raw ECMWF ENS rainfall forecasts. It is worth noting that what described to the following also applies to other NWP models. \parTo explain the reasons why ecPoint-Rainfall and raw ECMWF ENS rainfall forecasts may differ, let us consider an ENS grid cell and radar-derived totals for a rainfall event. Let us considered that the grid cell average rainfall total is about 17mm, while the minimum and maximum rainfall amounts are about 2mm and 60mm, respectively. The following can be stated. First, a completely accurate ENS member forecast would predict 17mm. However, this would not give the user any idea on what could be the spatial variability, within the grid cell, of the local (i.e. point) rainfall amounts when a particular G\_WT occurs. This describes an issue of sub-grid variability within the model grid box. Secondly, a particular G\_WT might show that an ENS member tends, on average, to over-predict (or under-predict) the grid-box rainfall by 15\%, delivering a forecast of 20mm (or 14mm) instead of 17mm. This describes an issue of model bias at grid-scale. By anticipating the sub-grid variability within the grid-box and correcting for biases in the model at grid-scale, ecPoint-Rainfall provides probabilities of bias-corrected rainfall forecasts at a point within the grid-box. However, ecPoint-Rainfall does not say where that point is within the ENS grid-box. \par
The calibration dataset for ecPoint-Rainfall is built applying the new concept of "Remote Calibration" described in \citet{Hewson2020a}. Typically, traditional post-processing methods use observations from a particular site. Such an approach limits the size of the calibration dataset, and forces to wait for at least ~20 years to have a robust calibration dataset, which still might not contain (enough) extremes cases for the considered site. The "Remote Calibration" approach removes any location labels and, instead, generates calibration datasets for particular G\_WTs pulling observations from where, around the world, those G\_WTs occurred. This approach facilitates the generation of extensive training datasets to post-process rainfall, including extremes. Indeed, the sites where a certain G\_WT occurs relatively often build the calibration dataset also for those sites where the same G\_WT would generate extreme rainfall. To give an order of magnitude, in order to create a similar calibration dataset for a location with similar size of one generated with the Remote Calibration approach, one would need to wait from hundreds to thousands of years. It is essential to highlight that the error bars on forecasts for globally extreme rainfall events (e.g. tropical cyclones) will be inevitably larger than for "not globally" extremes due to a lack of events in the calibration dataset. \par
 ecPoint-Rainfall outperforms raw ECMWF ENS in predicting localised rainfall events, including extremes \citet{Hewson2020a}. \par

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Fig0.png}}
\caption{Comparison of approaches for forecasts adjustments: expert human adjustment vs. ecPoint.}
\label{Fig.0}
\end{figure*}

\subsection{Hands-on experience, the trial}
 OMSZ and IMN volunteered to test the ecPoint-Rainfall forecasts in their day-to-day operational routines. The participation of OMSZ and IMN was ideal as the two meteorological services have different degrees of experience with probabilistic forecasts, and reside in two different climatological regions; tropical for IMN and extra-tropical for OMSZ. The trial period started in 2018. During the trial, OMSZ and IMN were invited to analyse the ecPoint-Rainfall performance in predicting extreme rainfall events. Given the small volumes of data available (only one year of data), both countries decided to base their analysis case studies. The selection of the case studies . The ecPoint-Rainfall forecasts were presented to both countries at the beginning of the trial. They were informed of different sources of information about ecPoint-Rainfall on the web, such as newsletters or the ECMWF Forecaster's User Guide \citep{Owens2018} and no further details were provided unless direct requests were submitted.

\subsection{Dialogue forecasters-developers}
The forecasters who monitored the ecPoint-Rainfall forecasts in their day-to-day operational duties were invited to discuss the outcomes of the reports. Forecasters were also invited to discuss the theoretical and practical challenges that emerged when adopting ecPoint-Rainfall in an operational environment. Such an open-ended approach is particularly well suited to explore behavioural responses to innovative forecast products like ecPoint-Rainfall, where the range of possible conceptual framings and the meaningful behaviour of the informants is not quantifiable but must be teased out through careful interpretation \citep{Patton2002}.

@@

\begin{figure}
\centerline{\includegraphics[width=18pc]{manuscript/Figures/Fig3.jpg}}
\caption{Product developed by IMN based on the 85th percentile of ecPoint-Rainfall.}
\label{Fig3}
\end{figure}

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Fig4.jpg}}
\caption{Panel (a) shows the climate regions of Costa Rica defined by IMN. Panels from (o-1) to (o-6) show 12-hourly rainfall observations between October 3rd and 5th, 2018. Panels from (i-1) to (i-3) show the impacts caused by the extreme event in the regions in which CNE declared the state of red alert (the photos come from www.teletica.com).}
\label{Fig4}
\end{figure*}

@@

\begin{figure}
\centerline{\includegraphics[width=19pc]{manuscript/Figures/Fig5.jpg}}
\caption{Panels (a) and (c) show the probabilities of not exceeding 50 mm/12h, and panels (b) and (d) show the 99th percentile, both for ecPoint-Rainfall forecasts. Panels (a) and (b) correspond to the forecast on 2018/10/04 at 00 UTC (t+18,t+30), which correspond to the period between 2018/10/04 12am and 201/10/05 0am (local time). Panels (c) and (d) correspond to the forecast on 2018/10/04 at 12 UTC (t+18,t+30), which correspond to the period between 2018/10/05 0am and 201/10/05 12am (local time).} 
\label{Fig5}
\end{figure}

IMN reported that the 85th percentile was selected to predict extreme events but to not focus only on the \textit{"worst-case scenarios"}. IMN forecasters pointed out, for example, that events around 50 mm/12h can already cause some impacts on the Pacific coast of Costa Rica or in the capital city, San José. Such events are relatively frequent in Costa Rica. Therefore, IMN forecasters thought they would miss such more frequent events if they look only at the highest percentiles. \par
IMN assessment is fair. On the other hand, by looking only at the 85th percentile (which would forecast events that have, on average, a 1 in 7 chance to occur), forecasters would miss the opportunity to be aware of what could be the possible "worst-case scenario". The latter becomes, indeed, a piece of vital information when preparing forecasts for critical decision-making processes in highly uncertain weather scenarios like the one analyzed. \par
During the dialogue with IMN forecasters, it was discussed the possibility to make available to forecasters the probabilities of exceeding a certain threshold. For example, 50 mm/12h is a rainfall event that can generate some impacts in some areas of Costa Rica. Forecasters could display on a map-plot the probabilities of exceeding 50 mm/12h (see mock-up product in Fig.\ref{Fig5}a and c) and identify whether in the concerning areas the probabilities of exceeding 50 mm/12h are higher than a probability threshold previously defined by IMN for which action needs to be taken. At the same time, by looking at the 99th percentile (see mock-up product in Fig.\ref{Fig5}b and d), forecasters could identify those areas at risk of being affected by the "worst-case scenario" (as it represents those events with, on average, a 1 in 100 chance to occur), and provide a richer warning on the areas at risk of highest impacts. \par
When both products (Fig.\ref{Fig5}) were shown to IMN forecasters forecasters to see whether they would have reformulated their forecasts for the case study, they produced a much more aware warning about the areas at risk of highest impacts (like those in https://www.imn.ac.cr/es/web/imn/avisos-meteorologicos). \par
Furthermore, subsequent examination of different extreme rainfall cases around the world has shown that raw ENS and ecPoint-Rainfall’s CDFs tend to cross around the 85th percentile, and ECMWF users might not extract extra information from ecPoint-Rainfall not provided by the raw ENS.

@@

\begin{figure*}
\centerline{\includegraphics[width=39pc]{manuscript/Figures/Fig6.jpg}}
\caption{Forecast evolution (from day 7 to day 1) for the rainfall event occurred between October 04th at 12am and October 5th at 0am (see Fig. \ref{Fig4}o-4 to compare the forecasts with the observations). From the left, the first column shows the 85th percentile for ecPoint-Rainfall, the second column shows the 85th percentile for the raw ENS, the third column shows the 99th percentile for ecPoint-Rainfall, the fourth column shows the wettest member of the raw ENS, and the fifth column shows the deterministic forecast from WRF-1.5 provided by IMN (spatial resolution of 1.5 km). The shown raw ECMWF ENS and ecPoint-Rainfall forecasts correspond to runs at 12 UTC, which is the first available run from Europe to IMN forecasters due to time difference (UTC-6) between Europe and America. The displayed WRF-1.5 forecasts correspond to runs at 18 UTC, which is the first run available to IMN forecasters in the morning for their in-house models.}
\label{Fig6}
\end{figure*}

@@

Fig.~\ref{Fig6} shows the forecasts, up to day 7, for the rainfall event between midday October 4th and midnight October 5th (compare with observations in Fig.\ref{Fig4}o-4). The first and the second column in Fig.~\ref{Fig6} show, respectively, the 85th percentile of ecPoint-Rainfall and raw ENS. The maps in these two columns show to the reader what it was previously stated about the similarity of the raw ENS and ecPoint-Rainfall's 85th percentile. The third, the fourth and the fifth column in Fig.~\ref{Fig6} show, respectively, the 99th percentile of ecPoint-Rainfall, the wettest ENS member, and the rainfall accumulations for the deterministic WRF-1.5 model. ecPoint-Rainfall provides from day 7 a more consistent guidance on the location of the areas at most risk of extreme rainfall, i.e. the whole Pacific coast (including the Nicoya peninsula), with a signal that becomes better defined from day 2 and shows forecast values of the same order as those observed. On the other hand, the raw ENS and the WRF-1.5 provide a noisier and jumpier signal. In the raw ENS the Nicoya peninsula is not highlighted in the guidance until day 2, and the rainfall extremes are underestimated of ~50\% also at day 1.

@@

%%%%%%%%%%%%%%%%%%%
% ACKNOWLEDGMENTS %
%%%%%%%%%%%%%%%%%%%
\acknowledgments
We thank the anonymous reviewers, Florian Pappenberger, David Richardson who helped to improve the manuscript with their comments.
